A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-14 EvoPar2017/paper.Rnw
Intro

\begin{abstract}
In a volunteer distributed computing system, users run a program on
their own machine to contribute to a common effort. If the program is
embedded in a web page, collaboration is straightforward, but also ephemeral. In
this paper, we analyze a volunteer evolutionary computing system called
NodIO, by running several experiments, some of them massive. Our
objective is to discover rules that encourage volunteer
participation and also the interplay of these contributions with the
dynamics of the algorithm itself, making it more or less efficient. We will show different measures of
participation and contribution to the algorithm and how different 
volunteer usage patterns and tweaks in the algorithm, such as
restarting clients when solution has been found, contribute to 
improvements and leveraging of these contributions. We will also try
to find out what is the key factor in the early termination of the
experiments, measuring entropy in the contributions and other large
scale indicators. 
\end{abstract}

@@

The World Wide Web provides not only a platform for content
distribution, but also, thanks to the maturity and reliability of the
HTTP protocol, an increasingly reliable and high-performance
operating system for running distributed applications. Besides the
protocol itself, there are two factors that contribute to this fact:
the Javascript virtual machine every browser runs \cite{paulson2005building}
and the simplified standard interface for interacting with servers
exemplified by the REST application interface convention \cite{masse2011rest}.
Thus creating a distributed computing experiment is just a matter of
making a JavaScript application interchange information with a server,
called REST. From the point of view of the programmer, this involves
relatively common skills and no special libraries, since that
interface is built in the browser, and a simple application that
responds to those requests on the server side; both involve just a few
dozens lines of code additionally to whatever business logic your
application has. But, more importantly and from the point of view of
the user, that application can be run by simply visiting a web page.

Using this system for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98}.

>>>>

\begin{abstract}
In a volunteer distributed computing system, users run a program on
their own machine to contribute to a common effort. If the program is
embedded in a web page, collaboration is straightforward, but also ephemeral. In
this paper, we analyze a volunteer evolutionary computing system called
NodIO, by running several experiments, some of them massive. Our
objective is to discover rules that encourage volunteer
participation and also the interplay of these contributions with the
dynamics of the algorithm itself, making it more or less efficient. We will show different measures of
participation and contribution to the algorithm and how different 
volunteer usage patterns and tweaks in the algorithm, such as
restarting clients when a solution has been found, contribute to 
improvements and leveraging of these contributions. We will also try
to find out what is the key factor in the early termination of the
experiments, measuring entropy in the contributions and other large
scale indicators. 
\end{abstract}

@@

The World Wide Web provides not only a platform for content
distribution, but also, thanks to the maturity and reliability of the
HTTP protocol, an increasingly reliable and high-performance
operating system for running distributed applications. Besides the
protocol itself, there are two factors that contribute to this fact:
the JavaScript virtual machine every browser runs \cite{paulson2005building}
and the simplified standard interface for interacting with servers
exemplified by the REST application interface convention \cite{masse2011rest}.
Thus creating a distributed computing experiment is just a matter of
making a JavaScript application interchange information with a server,
by using REST. From the point of view of the programmer, this involves
relatively common skills and no special libraries, since the
interface is built in the browser, and a simple application that
responds to those requests on the server side; both involve just a few
dozens lines of code additionally to whatever business logic the
application has. But, more importantly and from the point of view of
the user, that application can be run by simply visiting a web page.

Using this approach for creating distributed experiments is called {\em
  volunteer}, {\em cycle-scavenging} or {\em opportunistic} computing
\cite{sarmenta2001volunteer} and it dates back, in different shapes
and underlying mechanisms, to the origin of the web \cite{david-seti:home}. Our interest
here, however, is to use it as a resource for evolutionary
computation, as our group has done for a long time \cite{jj-ppsn98}.
